Fractional Types

### 2013-02-22 frac.tex
cleaned up currying stuff

We are now a small step from implementing various higher-order
combinators that manipulate functions or relations. In particular, we
can \emph{apply} and \emph{compose} values representing relations as
shown below:

@@

We can also derive currying (and dually, uncurrying) combinators. Observe that
the type of {{curry}} needs to be {{b1 * b2 :-* b3 <-> b1 :-* (b2 :-* b3)}},
which can be written in mathematical notation as {{b3/(b1 * b2) = (b3 * 1/b2) *
(1/b1)}}. This means {{curry}} can be written using {{recip}}, the
implementation of the mathematical identity {{1/(b1 * b2) = 1/b1 * 1/b2}}:

%subcode{opsem} include main
%! columnStyle = rcl
% recip &:& 1 / (b1 * b2) <-> 1 / b1 * 1 / b2
% recip &=& (uniti) (;) (uniti) (;) (assocl*) (;) ((eta* (*) eta*) (*) id)
%     &&    (;) (reorder (*) id) (;) (assocr*) (;) (id (*) swap*) (id (*) eps*)
%     &&    (;) swap* (;) (unite)

@@

%subcode{opsem} include main
%! columnstyle = rlr
%        & 1/(b1 * b2) & 
%     <->& 1 * 1/(b1 * b2) & uniti
%     <->& 1 * (1 * 1/(b1 * b2)) & uniti
%     <->& (1 * 1) * 1/(b1 * b2) & assocl*
%     <->& ((1/b1 * b1) * (1/b2 * b2)) * 1/(b1 * b2) &(eta*_{b1} (*) eta*_{b2}) (*) id
%     <->& ((1/b1 * 1/b2) * (b1 * b2)) * 1/(b1 * b2) & reoder (*) id
%     <->& (1/b1 * 1/b2) * ((b1 * b2) * 1/(b1 * b2)) & assocr*
%     <->& (1/b1 * 1/b2) * (1/(b1 * b2) * (b1 * b2)) & id * swap*
%     <->& (1/b1 * 1/b2) * 1 & id * eps*
%     <->& 1 * (1/b1 * 1/b2) & swap*
%     <->& (1/b1 * 1/b2) & unite

With {{recip}} out of the way, we can easily write {{curry}}:

@@

Given a computation {{c:b1 * b2 <-> b1 * b3}} and we
want to cancel the common type {{b1}}, to produce a new combinator
{{trace*~c:b2<->b3}}. For the evaluation of such a combinator, we are only
given a value {{v2 : b2}}. But to evaluate {{c}}, we must provide a pair
consisting of ``some'' value {{v1 : b1}} together with the given {{v2 : b2}}.
The needed value {{v1 : b1}} cannot be arbitrary: it must be such that is
is the value produced as the first component of the result.  In general, there
may be several such fixed-point values, or none.  A few examples should help
explain the subtleties.

>>>>

We are now a small step from implementing various higher-order
combinators that manipulate functions or relations. In particular, we
can \emph{apply} and \emph{compose} values representing relations.
\emph{apply} is shown below:

@@

\noindent
We can also derive currying (and dually, uncurrying) combinators. Observe that
the type of {{curry}} needs to be {{b1 * b2 :-* b3 <-> b1 :-* (b2 :-* b3)}},
which can be written in mathematical notation as {{b3/(b1 * b2) = (b3 * 1/b2) *
(1/b1)}}. This means {{curry}} can be written using {{recip}}, the
implementation of the mathematical identity {{1/(b1 * b2) = 1/b1 * 1/b2}}:

%subcode{opsem} include main
%! columnStyle = rcl
% recip &:& 1 / (b1 * b2) <-> 1 / b1 * 1 / b2
% recip &=& (uniti) (;) (uniti) (;) (assocl*) (;) ((eta*_{b1} (*) eta*_{b2}) (*) id)
%     &&    (;) (reorder (*) id) (;) (assocr*) (;) (id (*) swap*) (id (*) eps*)
%     &&    (;) swap* (;) (unite)

@@

%% XXX: why won't the columnstyle declaration here work? (Z)
%subcode{opsem} include main
%! columnStyle = rlr
%        & 1/(b1 * b2) & 
%     <->& 1 * 1/(b1 * b2) & (uniti)
%     <->& 1 * (1 * 1/(b1 * b2)) & (uniti)
%     <->& (1 * 1) * 1/(b1 * b2) & (assocl*)
%     <->& ((1/b1 * b1) * (1/b2 * b2)) * 1/(b1 * b2) &((eta*_{b1} (*) eta*_{b2}) (*) id)
%     <->& ((1/b1 * 1/b2) * (b1 * b2)) * 1/(b1 * b2) & (reoder (*) id)
%     <->& (1/b1 * 1/b2) * ((b1 * b2) * 1/(b1 * b2)) & (assocr*)
%     <->& (1/b1 * 1/b2) * (1/(b1 * b2) * (b1 * b2)) & (id * swap*)
%     <->& (1/b1 * 1/b2) * 1 & (id * eps*)
%     <->& 1 * (1/b1 * 1/b2) & (swap*)
%     <->& (1/b1 * 1/b2) & (unite)

\noindent
With {{recip}} out of the way, we can easily write {{curry}}:

@@

Given a computation {{c:b1 * b2 <-> b1 * b3}},we
want to cancel the common type {{b1}}, to produce a new combinator
{{trace*~c:b2<->b3}}. For the evaluation of such a combinator, we are only
given a value {{v2 : b2}}. But to evaluate {{c}}, we must provide a pair
consisting of ``some'' value {{v1 : b1}} together with the given {{v2 : b2}}.
The needed value {{v1 : b1}} cannot be arbitrary: it must be such that it
is the value produced as the first component of the result.  In general, there
may be several such fixed-point values, or none.  A few examples should help
explain the subtleties.
